Log-likelihood for summary data

### 2023-01-23 Working Rmds/summary_loglike_lognormal.Rmd
Correctly log-transform predictions if log = TRUE specified when computing RMSE, R-squared, etc.
Update documentation.

>>>>

# Appendix: Error in log-scale moments from arithmetic moments, a simulation study

How much error is there in estimating the log-scale moments (mean and SD of log-transformed observations) from the arithmetic moments (mean and SD of natural-scale observations)?

We can get an idea using a simulation study.

Take a log-normal distribution with known log-scale mean and SD.

Draw some number of samples from this distribution.

Calculate the sample mean and SD of the log-scaled samples, and of the natural-scale samples.

Repeat these steps some large number of times.

Compare the log-scale sample means and SDs to the natural-scale sample means and SDs.

Write a function to draw the samples and calculate the sample moments:

```{r}
get_sample_moments <- function(meanlog, sdlog, n_samp){
  #samples
  y <- rlnorm(n = n_samp, meanlog = meanlog, sdlog = sdlog)
  #log-transformed samples
  ylog <- log(y)
  return(c("sample_mean" = mean(y),
           "sample_sd" = sd(y),
           "sample_meanlog" = mean(ylog),
           "sample_sdlog" = sd(ylog),
           "sample_n" = n_samp))
}
```

Choose values for meanlog, sdlog, and n_samp
```{r}
meanlog <- 0
sdlog <- 1
n_samp <- 7
```

Set a seed for reproducibility, and generate the values

```{r}
set.seed(42)
n_rep <- 1000
samp <- replicate(n = n_rep,
          expr = get_sample_moments(meanlog = meanlog,
                                    sdlog = sdlog,
                                    n_samp = n_samp))
samp <- t(samp)
samp <- as.data.frame(samp)
```

Calculate the estimated log-scale sample moments from the arithmetic sample moments

```{r}
samp[c("est_sample_meanlog",
       "est_sample_sdlog")] <- convert_summary_to_log(sample_mean = samp$sample_mean,
                                                                   sample_SD = samp$sample_sd)
```

Now, plot the estimated log-scale moments vs. the "actual" log-scale moments.

Log-scale sample mean. The "true" population log-scale mean (0.0) is indicated by vertical and horizontal dashed lines.

```{r}
ggplot(data = samp) +
  geom_point(aes(x = sample_meanlog,
                 y = est_sample_meanlog),
             color = "gray50") +
  geom_abline(intercept = 0, slope = 1, color = "black") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  xlab("Actual log-scale sample mean") +
  ylab("Estimated log-scale sample mean") +
  ggtitle("Estimated vs. actual log-scale sample mean, n = 7") +
  theme_bw()
```

The estimated log-scale sample mean usually overestimates the "true" log-scale sample mean (most points are above the identity line).

Log-scale sample SD. The "true" population log-scale SD (1.0) is indicated by vertical and horizontal dashed lines.

```{r}
ggplot(data = samp) +
  geom_point(aes(x = sample_sdlog,
                 y = est_sample_sdlog),
             color = "gray50") +
  geom_abline(intercept = 0, slope = 1, color = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
  geom_vline(xintercept = 1, linetype = 2) +
  xlab("Actual log-scale sample SD") +
  ylab("Estimated log-scale sample SD") +
  ggtitle("Estimated vs. actual log-scale sample SD, n = 7") +
  theme_bw()
```

The estimated log-scale sample SD usually underestimates the actual log-scale sample SD (most points are below the identity line).

```{r}
summary(lm(est_sample_sdlog ~ sample_sdlog, data = samp))
```
